---
title: "Part1"
author: "Ugne Milasiunaite"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Task 1

We load functions from the file.

```{r, code = readLines("funktioner.R")}
```

Now, we load the data and configure the model.

```{r}
load("proj_data.Rdata")
modell <- glm(Resultat ~ Alder + Kon + Utbildare, 
              data = data_individ,
              family = "binomial")
(model_output <- summary(modell))
```

Next, we configure X and y.

```{r}
y <- matrix(data_individ$Resultat, ncol = 1)
X <- model.matrix(Resultat ~ Alder + Kon + Utbildare, data = data_individ)
```

## Task 2

We verify our implementation of Newton-Rhapson's algorithm by computing parameter estimates.

```{r}
theta0 <- c(0, 0, 0, 0)
```

## Task 3

We compute the approximations of standard errors by taking the square root of the diagonal elements of Fisher Information.

```{r}
errors <- sqrt(diag(solve(I(modell$coefficients, y, X))))
abs(errors - model_output$coefficients[, 2])
```

Seems like the differences are very small.

## Task 4

Approximate the errors using parametric Bootstrap.

```{r}

```
