---
title: "Part2"
author: "Ugne Milasiunaite (20010905-8420) & Meike Knull (20001220-T588)"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We load the functions from Part1.

```{r}
source("funktioner.R")
```

Next, we recreate the model from previous task.

```{r}
load("proj_data.Rdata")
modell <- glm(Resultat ~ Alder + Kon + Utbildare,
              data = data_individ,
              family = "binomial")
(model_output <- summary(modell))
y <- matrix(data_individ$Resultat, ncol = 1)
X <- model.matrix(Resultat ~ Alder + Kon + Utbildare,
data = data_individ)
```

## Task 1

We verify, using functions I and NR from Part I, that the z value column of the output are Wald statistics. Note that z values in the output test if the true parameter is 0.

```{r}
zval <- model_output$coefficients[, 3]
theta0 <- c(0, 0, 0, 0)
niter <- 100
mle_theta <- NR(theta0, niter, y, X)
se <- sqrt(diag(solve(I(mle_theta, y, X))))
abs(mle_theta/se - zval)
```

We see that the difference between our computed statistic and the output of R's summary function for z values are identical for 5 significant digits for all parameter values.

## Task 2

We compute the generalized likelihood ratio statistics that correspond to the Wald statistics in Task 1 and determine the corresponding p-values.

```{r}
coefs <- model_output$coefficients[, 1]

# statistic for intercept
eta <- NR(theta0 = c(0, 0, 0), niter = 10, y = y, X = X[, -1])
W1 <- 2 * log(L(coefs, y, X)/L(c(0, eta), y, X))

# statistic for age
eta <- NR(theta0 = c(0, 0, 0), niter = 10, y = y, X = X[, -2])
W2 <- 2 * log(L(coefs, y, X)/L(c(eta[1], 0, eta[2:3]), y, X))

# statistic for sex
eta <- NR(theta0 = c(0, 0, 0), niter = 10, y = y, X = X[, -3])
W3 <- 2 * log(L(coefs, y, X)/L(c(eta[1:2], 0, eta[3]), y, X))

# statistic for education
eta <- NR(theta0 = c(0, 0, 0), niter = 10, y = y, X = X[, -4])
W4 <- 2 * log(L(coefs, y, X)/L(c(eta, 0), y, X))

```

Check that we have the same order of magnitude.

```{r}
matrix(c(W1, W2, W3, W4, as.vector(zval) ** 2), ncol = 4, byrow = TRUE, 
       dimnames = list(c("W", "z^2"), names(zval)))
```

Now calculate the p-values.

```{r}

```

## Task 3

We compute the ML estimate under the null hypothesis and use the result to determine a p-value based on the generalized score statistic.

```{r}
# ml estimate
estm <- NR(theta0 = c(0, 0), niter = 10, y = y, X = X[, 3:4])
# parameter values under null hypothesis
theta_h0 <- c(0, 0, estm)
# generalized score test statistic
score <- S(theta_h0, y, X)
Ts <- t(score) %*% solve(I(theta_h0, y, X)) %*% score
# calculate p-value
```

## Task 4

We compute the profile likelihood for parameter Q on a grid of parameter values XXX.

```{r}
grid <- seq(-0.3, 0.3, 0.02)
n_thetas <- length(grid)
Lps <- numeric(n_thetas)
Les <- numeric(n_thetas)
for (i in 1:n_thetas) {
  profil <- glm.fit(x = X[, -3], y = y, 
                    offset = grid[i] * X[, 3],
                    family = binomial())
  coefs_h0 <- profil$coefficients
  Lps[i] <- L(c(coefs_h0[1:2], grid[i], coefs_h0[3]), y, X)
  Les[i] <- L(c(coefs[1:2], grid[i], coefs[4]), y, X)
}
```

Now, we choose the level for confidence interval.

```{r}
# construct a test value for profile likelihood
Lp_test <- L(coefs, y, X) * exp(-1/2 * qchisq(p=0.95, df=1))
```

We use the results to graph the profile likelihood Lp together with the corresponding estimated likelihood Le and the line signifying the threshold for 95% significance.

```{r}
plot(grid, Lps, pch = 20, ylim = c(0, max(Lps)), ylab = "Likelihood", xaxp = c(-0.3, 0.3, 12))
points(grid, Les, pch = 3)
abline(h = Lp_test, lty = "dashed")
legend(x = "topleft", legend = c("95% threshold", "Estimated", "Profile"), pch = c(NA, 3, 20), lty = c("dashed", NA, NA), bty = "n")
```

Now we compare the intervals.

For profile likelihood, we have that the 95% confidence interval is $\{L_p(\theta_{Kon}) \ge L_p(\hat\theta_{ML})\exp(-0.5\chi_{0.95}^2(1))\} = \{L_p (\theta_{Kon}) \ge Lptest\}$. For Wald interval, the z value is `r qnorm(p=0.975)`.

```{r}
lbound <- coefs[3] - qnorm(p=0.975) * model_output$coefficients[3, 2]
ubound <- coefs[3] + qnorm(p=0.975) * model_output$coefficients[3, 2]
```

So the Wald interval is [`r lbound`, `r ubound`]. Compare visually.

```{r}
plot(grid, Lps, type = "l", ylim = c(0, max(Lps)), ylab = "Likelihood", xaxp = c(-0.3, 0.3, 12))
abline(h = Lp_test, lty = "dashed")
abline(v = lbound, col = "red")
abline(v = ubound, col = "red")
legend(x = "topleft", legend = c("threshold Lp", "Profile", "bounds Wald"), lty = c("dashed", "solid", "solid"), col = c("black", "black", "red"), bg = "white")
```

